---
title: "CVHH_FLASH-seq"
author: "Amalie Grenov"
format: html
editor: visual
---

## CVHH FLASH-seq analysis

Load R packages

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(biomaRt)
```

Create object for each plate

```{r}
read_plate <- function(n) {
  file <- sprintf("nfCore_run/Plate%d/plate%d-results/star_rsem/rsem.merged.gene_counts.tsv", n, n)
  data <- read.delim(file, row.names = 1)
  as.matrix(data[, -1])
}

setwd("Z:/home/users/grenova/K-tag/FLASH-seq")
Plate1.matrix <- read_plate(1) # Plasma cells, GFP and Tom CVHH4.1g
Plate2.matrix <- read_plate(2) #Plasma cells, GFP and Tom CVHH3.1d
Plate3.matrix <- read_plate(3) #ABCs, GFP and Tom CVHH3.1d and CVHH4.1g
Plate4.matrix <- read_plate(4) #BM plasma cells. GFP and Tom. CVHH4.1g and CVHH3.1d
Plate5.matrix <- read_plate(5) #BM and Spleen plasma cells. GFP and Tom. CVHH3.1g

```

Initialize the Seurat objects with the raw (non-normalized data).

```{r}
make_seurat <- function(matrix, project) {
  obj <- CreateSeuratObject(counts = matrix, project = project, min.cells = 3, min.features = 200)
  return(obj)
}

Plate1 <- make_seurat(Plate1.matrix, "Plate1")
Plate2 <- make_seurat(Plate2.matrix, "Plate2")
Plate3 <- make_seurat(Plate3.matrix, "Plate3")
Plate4 <- make_seurat(Plate4.matrix, "Plate4")
Plate5 <- make_seurat(Plate5.matrix, "Plate5")

```

Generate violin plots for all plates

```{r}
# Put all matrices in a named list
Plates <- list(Plate1, Plate2, Plate3, Plate4, Plate5)
names(Plates) <- c("Plate1", "Plate2", "Plate3", "Plate4", "Plate5")

# Generate violin plots for all plates
vln_plots <- lapply(names(Plates), function(name) {
  VlnPlot(Plates[[name]], features = c("nCount_RNA", "nFeature_RNA")) +
    ggtitle(name)
})

# Combine and print all violin plots together
wrap_plots(vln_plots)
```

# **Add mitochondrial gene counts**

Retrieve Mt gene ENSMEBL identifiers from biormart

```{r}
# Get mitochondrial gene ENSEMBL IDs for mouse
mart <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")

# Check which filters exist (you can run listFilters(mart) if curious)
# We’ll filter by chromosome name instead, since mitochondrial genes are on "MT"

mt_genes <- getBM(
  attributes = c("ensembl_gene_id", "chromosome_name"),
  filters = "chromosome_name",
  values = "MT",
  mart = mart
)$ensembl_gene_id

length(mt_genes)
head(mt_genes)

```

Then calculate % mitochondrial genes per Seurat object

```{r}

for (i in seq_along(Plates)) {
  obj <- Plates[[i]]
  obj[["percent.mt"]] <- PercentageFeatureSet(obj, features = mt_genes)
  Plates[[i]] <- obj
}

# Optional: reassign back
Plate1 <- Plates$Plate1
Plate2 <- Plates$Plate2
Plate3 <- Plates$Plate3
Plate4 <- Plates$Plate4
Plate5 <- Plates$Plate5
```

QC visualization

```{r}
wrap_plots(lapply(names(Plates), function(name) {
  VlnPlot(Plates[[name]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt")) + ggtitle(name)
}))

setwd("Z:/home/users/grenova/K-tag/FLASH-seq/Exported_Figures")
ggsave("QC_visualization.png", plot = last_plot(), width = 20, height = 8, dpi = 300)
```

Subset Objects to include only cells with more than limit nFeature_RNA and percent_mt

```{r}
head(Plate1@meta.data)

# Filter all at once
Plates <- lapply(Plates, function(x) {
  subset(x, subset = nFeature_RNA > 200 & percent.mt < 15)
})

# Assign back to individual objects (optional)
Plate1 <- Plates$Plate1
Plate2 <- Plates$Plate2
Plate3 <- Plates$Plate3
Plate4 <- Plates$Plate4
Plate5 <- Plates$Plate5

```

### Add Meta data

First Add Cellnumber column to Sample info sheet\
NB: SeqID number corresponds to Cell# (See nfCore_run_Samplesheet_preparation.R)

```{r}
setwd("Z:/home/users/grenova/K-tag/FLASH-seq")
Sample_Info <- read.csv("SC24265_Sample_info3.csv")

Sample_Info$cell <- paste0(
  "cell",
  sub(".*?(\\d{1,4})$", "\\1", Sample_Info$Sample.limsid)
)

rownames(Sample_Info) <- Sample_Info$cell
```

Add metadata for cell type and fate_map (GFP/Tomato)

```{r}

Plate1 <- AddMetaData(Plate1, metadata = Sample_Info)
Plate2 <- AddMetaData(Plate2, metadata = Sample_Info)
Plate3 <- AddMetaData(Plate3, metadata = Sample_Info)
Plate4 <- AddMetaData(Plate4, metadata = Sample_Info)
Plate5 <- AddMetaData(Plate5, metadata = Sample_Info)


head(Plate2@meta.data)
```

### Add metadata for BCR alignment

```{r}
setwd("Z:/home/users/grenova/K-tag/FLASH-seq/BraCeR")
BCR_align <- read.csv("IMGT_gapped_db_Mouse_all_reads.csv")

BCR_align$cell <- paste0(
  "cell",
  sub(".*?(\\d{1,4})$", "\\1", BCR_align$CELL)
)

# Subset so rows with "H" in LOCUS come first
BCR_align_H <- subset(BCR_align, grepl("H", LOCUS))

head(BCR_align_H)

# Count duplicates

sum(duplicated(BCR_align_H$cell))
```

**remove rows where `"cell"` is duplicated and `"C_GENE"` is `"None"`**, keeping at least one of each duplicate if it has a real `C_GENE` value.

```{r}
BCR_align_H_clean <- BCR_align_H %>%
  group_by(cell) %>%
  filter(!(C_GENE == "None" & n() > 1)) %>%
  ungroup()

sum(duplicated(BCR_align_H_clean$cell))
```

Got it! You want to **remove rows with duplicates in `"cell"` where the `"V_EVALUE"` is the highest**, keeping the row with the lowest (best) `V_EVALUE` for each `"cell"`. Here's a clean way to do it using **`dplyr`** in R

```{r}
BCR_align_H_clean <- BCR_align_H_clean %>%
  group_by(cell) %>%
  slice_min(order_by = V_EVALUE, n = 1) %>%  # Keep row(s) with smallest V_EVALUE
  ungroup() %>%
  distinct(cell, .keep_all = TRUE)           # Remove any leftover duplicates
sum(duplicated(BCR_align_H_clean$cell))
```

Now can asign cell to row.names

```{r}
rownames(BCR_align_H_clean) <- BCR_align_H_clean$cell
```

Add BCR_alignment data to Seurat object

```{r}
colnames(BCR_align_H_clean)

#Save rownames as column
BCR_align_H_clean <- BCR_align_H_clean %>% 
  tibble::rownames_to_column(var = "cell_id")  # create a column with the row names

BCR_H_asign <- BCR_align_H_clean %>%
  dplyr::select(cell_id, FUNCTIONAL, V_CALL, D_CALL, J_CALL, ISOTYPE, CLONE)

#convert to dataframe
BCR_H_asign <- as.data.frame(BCR_H_asign)

rownames(BCR_H_asign) <- BCR_H_asign$cell
BCR_H_asign$cell_id <- NULL
```

Now finally add BCR data to Meta Data

```{r}
Plate1 <- AddMetaData(Plate1, metadata = BCR_H_asign)
Plate2 <- AddMetaData(Plate2, metadata = BCR_H_asign)
Plate3 <- AddMetaData(Plate3, metadata = BCR_H_asign)
Plate4 <- AddMetaData(Plate4, metadata = BCR_H_asign)
Plate5 <- AddMetaData(Plate5, metadata = BCR_H_asign)

head(Plate2@meta.data)
```

### Merge objects:

```{r}
CVHHseq <- merge(x = Plate1, y = c(Plate2, Plate3, Plate4, Plate5), add.cell.ids = c("Plate1", "Plate2", "Plate3", "Plate4", "Plate5"), project = "CVHHseq")
table(CVHHseq$orig.ident)
VlnPlot(CVHHseq, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"))
```

### Process reads, normalize, cluster, dimensionality reduction

Normalize data

First Regress out immunoglobulin genes \# NB remove also IgK and IgL genes

```{r}

all_genes <- getBM(
  attributes = c("ensembl_gene_id", "external_gene_name"),
  mart = mart
)

igh_genes <- all_genes$ensembl_gene_id[grep("^Igh", all_genes$external_gene_name)]
igl_genes <- all_genes$ensembl_gene_id[grep("^Igl", all_genes$external_gene_name)]
igk_genes <- all_genes$ensembl_gene_id[grep("^Igk", all_genes$external_gene_name)]

length(igh_genes)
head(igh_genes)

ig_genes <- c(igh_genes, igl_genes, igk_genes)

length(ig_genes)

#Compute Ig score
CVHHseq$IG_score <- PercentageFeatureSet(CVHHseq, features = ig_genes)
```

Normalize using SCT

```{r}
CVHHseq <- SCTransform(CVHHseq, vars.to.regress = "IG_score")

#| echo: false
```

Run PCA - determine number of dimensions

```{r}
CVHHseq <- RunPCA(CVHHseq, features = VariableFeatures(object = CVHHseq))

# Examine and visualize PCA results a few different ways
print(CVHHseq[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(CVHHseq, dims = 1:2, reduction = "pca")
DimPlot(CVHHseq, reduction = "pca") + NoLegend()
DimHeatmap(CVHHseq, dims = 1:10, balanced = TRUE)
ElbowPlot(CVHHseq)
```

Run UMAP

```{r}
CVHHseq <- FindNeighbors(CVHHseq, dims = 1:10)
CVHHseq <- FindClusters(CVHHseq, resolution = 0.7)
CVHHseq <- RunUMAP(CVHHseq, dims = 1:10)
```

Save RDS file

```{r}
setwd("Z:/home/users/grenova/K-tag/FLASH-seq/R_scripts")
saveRDS(CVHHseq, file = "251024CVHHseq.rds")
setwd("Z:/home/users/grenova/K-tag/FLASH-seq/R_scripts")
CVHHseq <- readRDS("251024CVHHseq.rds")
```

Visualize UMAP

```{r}
My12Colours <- c('0'="#E7B4BC",'5'="#D8E9E6", '4'="#CACDDC", '3'="#F2A843", '2'="#979FC6", '11'="#EAC4AF", '10'="#5BD285", '1'="#E694A6", '9'="#979797", '6'="#7791EA", '7'="#AFE4DE", '8'="#9A3E67" )

head(CVHHseq@meta.data)
Idents(CVHHseq) <- "seurat_clusters"
DimPlot(CVHHseq, reduction = "umap", pt.size = 2, cols = My12Colours)
DimPlot(CVHHseq, split.by = "Cell.Type", pt.size = 2, cols = My12Colours)

setwd("Z:/home/users/grenova/K-tag/FLASH-seq/Exported_Figures")
ggsave("umap.png", plot = last_plot(), width = 10, height = 8, dpi = 300)

setwd("Z:/home/users/grenova/K-tag/FLASH-seq/Exported_Figures")
ggsave("umap_orig_ident.png", plot = last_plot(), width = 18, height = 8, dpi = 300)
```

Rename Seurat clusters

```{r}
# Example: rename clusters 0–3
new.cluster.ids <- c(
  "0" = "IgM PCs",
  "1" = "CS PCs",
  "2" = "IgM ABC/PCs",
  "3" = "IgA PCs",
  "4" = "Spleen IgM PCs",
  "5" = "ABCs",
  "6" = "IgD ABCs",
  "7" = "ABC/PB",
  "8" = "IgD ABC/PCs"
)

# Apply new names
CVHHseq <- RenameIdents(CVHHseq, new.cluster.ids)
```

Add cell cycle score

```{r}
# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

CVHHseq <- CellCycleScoring(CVHHseq, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

# view cell cycle scores and phase assignments
head(combined_object[[]])

DimPlot(combined_object, reduction = "umap")
```

Add ABC signature

Highlight clones

Each colour different, NA as gray

```{r}

# Define the clones you want to highlight
highlight_clones <- c("375", "526", "652", "567")

# Create a new column for plotting
CVHHseq$clone_plot <- ifelse(CVHHseq$CLONE %in% highlight_clones, 
                             CVHHseq$CLONE, "Other")

# Define colors: one color per highlighted clone, plus gray for all others
library(RColorBrewer)
n_colors <- length(highlight_clones)
colors <- c(RColorBrewer::brewer.pal(n = n_colors, name = "Set3"), "gray")  # last color is gray for "Other"


# Manually define colors for each highlighted clone
manual_colors <- c(
  "375" = "#9A3E67",  # red
  "526" = "#377EB8",  # blue
  "652" = "#5BD285",  # green
  "567" = "#E694A6",  # purple
  "92" = "#7791EA",  # orange
  "188" = "#EAC4AF",  # yellow
  "35" = "#AFE4DE",  # brown
  "71" = "#E7B4BC",  # pink
  "00" = "#979FC6", # gray-ish (or pick another bright color)
  "181" = "#66b9bd", # teal
  "Other" = "#CACDDC"  # all other cells, including NAs
)

# Plot
DimPlot(CVHHseq, group.by = "clone_plot", cols = colors, pt.size = 2)
DimPlot(CVHHseq, group.by = "clone_plot", cols = manual_colors, pt.size = 2)

setwd("Z:/home/users/grenova/K-tag/FLASH-seq/Exported_Figures")
ggsave("clone_plot.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

**Colour dimplot by isotype**

```{r}
# Replace NA with "Other"
CVHHseq$ISOTYPE <- ifelse(is.na(CVHHseq$ISOTYPE), "Other", CVHHseq$ISOTYPE)

# First, ensure ISOTYPE has "None", "Other", "Unknown"
# Move these to the front (background), others stay on top
iso_levels <- c("None", "Other", "Unknown", 
                setdiff(unique(CVHHseq$ISOTYPE), c("None","Other","Unknown")))

CVHHseq$ISOTYPE <- factor(CVHHseq$ISOTYPE, levels = iso_levels)


# Manually define colors for each isotype
isotype_colors <- c(
  "IGHA" = "lightgreen",  
  "IGHG2B" = "#1b9e77",  
  "IGHG2A" = "#bf5b17",
  "IGHG3" = "#beaed4",  
  "IGHG2C" = "#386cb0",
  "IGHG1" = "#EAC4AF",
  "IGHM" = "#666666", 
  "IGHD" = "#E7B4BC", 
  "None" = "#CACDDC", # gray-ish (or pick another bright color)
  "Unknown" = "#CACDDC", # teal
  "Other" = "#CACDDC"  # all other cells, including NAs
)


# Basic DimPlot colored by ISOTYPE
DimPlot(CVHHseq, group.by = "ISOTYPE", pt.size = 2, cols = isotype_colors) 
  

setwd("Z:/home/users/grenova/K-tag/FLASH-seq/Exported_Figures")
ggsave("Isotype_umap.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
#| echo: false
```

Feature plots

```{r}
FeaturePlot(CVHHseq, features = c(""))

FeaturePlot(CVHHseq, features = "ENSMUSG00000038151", pt.size = 2) # Prdm1  
FeaturePlot(CVHHseq, features = "ENSMUSG00000067149", pt.size = 2) # Jchain
FeaturePlot(CVHHseq, features = "ENSMUSG00000020484", pt.size = 2) # Xbp1
FeaturePlot(CVHHseq, features = "ENSMUSG00000022496", pt.size = 2) # BCMA tnfrsf17
FeaturePlot(CVHHseq, features = "ENSMUSG00000020592", pt.size = 2) # Sdc1

FeaturePlot(CVHHseq, features = "ENSMUSG00000030789", pt.size = 2) # Itgax/Cd11c
FeaturePlot(CVHHseq, features = "ENSMUSG00000001444", pt.size = 2) # Tbx21/T-bet
FeaturePlot(CVHHseq, features = "ENSMUSG00000026872", pt.size = 2) # Zeb2
FeaturePlot(CVHHseq, features = "ENSMUSG00000048031", pt.size = 2) # FCRL5
FeaturePlot(CVHHseq, features = "ENSMUSG00000036469", pt.size = 2) # March1 

FeaturePlot(CVHHseq, features = "ENSMUSG00000033306", pt.size = 2, label = T) # Lpp
setwd("Z:/home/users/grenova/K-tag/FLASH-seq/Exported_Figures")
ggsave("Lpp_expression.png", plot = last_plot())

FeaturePlot(CVHHseq, features = "ENSMUSG00000024669", pt.size = 2) # Cd5 
FeaturePlot(CVHHseq, features = "ENSMUSG00000005540", pt.size = 2) # Fcer2a / CD23 

FeaturePlot(CVHHseq, features = "ENSMUSG00000076617", pt.size = 2) # Ighm
FeaturePlot(CVHHseq, features = "ENSMUSG00000104213", pt.size = 2) # Ighd
FeaturePlot(CVHHseq, features = "ENSMUSG00000076612", pt.size = 2) # Ighg2c
FeaturePlot(CVHHseq, features = "ENSMUSG00000076613", pt.size = 2) # Ighg2b
FeaturePlot(CVHHseq, features = "ENSMUSG00000076614", pt.size = 2) # Ighg1
FeaturePlot(CVHHseq, features = "ENSMUSG00000095079", pt.size = 2) # Igha

FeaturePlot(CVHHseq, features = c("ENSMUSG00000076617","ENSMUSG00000104213", "ENSMUSG00000076612", "ENSMUSG00000076613", "ENSMUSG00000076614", "ENSMUSG00000095079" ) , pt.size = 2) # Igha

FeaturePlot(CVHHseq, features = "ENSMUSG00000020914", pt.size = 2) # Top2a
FeaturePlot(CVHHseq, features = "ENSMUSG00000031004", pt.size = 2) # Mki67
FeaturePlot(CVHHseq, features = "ENSMUSG00000022346", pt.size = 2) # Myc
```
